---
title: "Neural Based Inference for the Browian Motion Model"
output:
  pdf_document: default
  html_document: default
date: "2025-09-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# I - Packages et fonctions

```{r,results='hide'}
library(reticulate)
use_virtualenv("~/r-tf", required = TRUE) 
library(tensorflow)
library(keras)
library(simrec)
library(openxlsx)
library(scales)
library(ggplot2)
library(gridExtra)
py_config() 
library(tidyr)
library(dplyr)
library(ggnewscale)
```

```{r}
detect_records_maxima_norm<-function(X){
  l=length(X)
  rec_max=c(X[1])
  ind_max<-c(1)
  
  for(i in 2:(l-1)){
    #print(i)
    if(X[i-1]<=X[i] && X[i]>X[i+1]){
      l<-length(rec_max)
      if(X[i]> rec_max[l]){
        rec_max<-c(rec_max,X[i])
        ind_max<-c(ind_max,i)
      }
    }
  }
  df<-data.frame(rec_max,ind_max)
  return(df)
}
```

# II - Préparation des échantillons


## 1 - Paramétrisation

### a) Utilisation des données du glacier des Bossons
```{r}
bossons <- read.xlsx("/home/mmegret/Documents/thèse Maud MEGRET/données/bossons2Phil.xlsx",sep=';')
bossons_0_NA<-bossons[-which(bossons$`Position;front;(mean)`==NaN),]
year<-bossons_0_NA$Year
year_NA<-bossons$Year
# en utilisant la bonne représentation 

pos_inv<-rev(bossons_0_NA$`Position;front;(mean)`)
pos_min_inv<-rev(bossons_0_NA$`Position;front;(min)`)
pos_max_inv<-rev(bossons_0_NA$`Position;front;(max)`)

pos_inv_NA<-rev(bossons$`Position;front;(mean)`)
pos_min_inv_NA<-rev(bossons$`Position;front;(min)`)
pos_max_inv_NA<-rev(bossons$`Position;front;(max)`)

pos_NA<--as.numeric((bossons$`Position;front;(mean)`))
pos_min_NA<--as.numeric((bossons$`Position;front;(min)`))
pos_max_NA<--as.numeric((bossons$`Position;front;(max)`))

X_NA<-as.numeric(pos_inv_NA)+1330
X_0<-X_NA
X_0[which(is.na(X_0))]<-0
l_1_0<-(detect_records_maxima_norm(X_0)$ind_max[-1])[1]
l_ind_0<-detect_records_maxima_norm(X_0)$ind_max[-1]
recordTimes_0=l_ind_0
recordValues_0=(detect_records_maxima_norm(X_0)$rec_max[-1])
dMor_0=cbind(recordTimes_0,recordValues_0)
dMor_0<-dMor_0[-4,]
dMor_0<-dMor_0[-c(6,7,9),]

mor1_0=dMor_0[1,1]
morf_0=dMor_0[length(dMor_0[,1]),1]

years_mor<-(mor1_0:morf_0)[dMor_0[,1]]
w_1818<-which(year_NA>=1818)
w_18_99<-which((year_NA[w_1818])<=1999)

pos_0_NA<-pos_NA[w_1818]
pos_0_NA[which(is.na(pos_0_NA))]<-0

```


### b) Graphique

Transformation of the observed \textit{Glacier des Bossons} trajectory from 1818 to 1999 to fit the record theory framework. Source: \cite{nussbaumer2012little}. The blue dots are the observed moraines, and the black line is the real observed trajectory of the frontal position of the \textit{Glacier des Bossons}. Both are measured along a longitudial axis corresponding to the glacier flow direction. Subfigure a) : the x-axis represents the years and the y-axis the frontal position relative to the first moraine position that corresponds to the glacier maximal extent during the Littel Ice Age. Subfigure b) : the x-axis represents the number of years between 1999 and the year plotted and the y-axis position relative to the 1999 glacier frontal position.

```{r}
par(mfrow=(c(1,2)))
plot((year_NA[w_1818])[w_18_99],-(pos_NA[w_1818])[w_18_99],type='l',xlab='Time (year)',ylab='Cumulative length change (m)',lwd=2)
points(2011-dMor_0[,1],dMor_0[,2]-1330,pch=19,cex=2,col='blue')
plot((mor1_0:morf_0)-12,X_NA[mor1_0:morf_0],type='l',xlab='Time from the last moraine (year)',ylab='Relavtive position (m)',lwd=2)
points(dMor_0[,1]-12,dMor_0[,2],pch=19,cex=2,col='blue')
legend('bottomright',legend=c('Moraines','Collected data'),pch=c(19,NA),lwd=c(NA,2),col=c('blue','black'))
```


### c) Paramètres

```{r}
N=length(mor1_0:morf_0)  #taille des trajectoires
n_samples <- 100000 # nombre d'échantillon

# matrice des trajectoires
X_bm_NA <- matrix(NA, nrow = n_samples, ncol = N) 
# matrice des paramètres du modèle brownien
params_bm_NA <- matrix(0, nrow = n_samples, ncol = 2) 

SigmaG <-1 #fixé
Ssmin<-0
Ssmax<-10
Sdmin<-50
Sdmax<-200
```

### d) Remplissage des matrices de trajectoires et de paramètres associés

Code non executé car données sauvegardées

```{r, eval=FALSE}
for (i in 1:n_samples) {
  SigmaS <- runif(1, Ssmin,Ssmax)
  Sd <- runif(1, Sdmin,Sdmax)
  #print(i)

  traj_sim <- simRecords(dMor_0,
                         N=N,
                         gaussConv = TRUE,
                         sigmaGauss = SigmaG,
                         sigmaSmooth=SigmaS,
                         sd = Sd)[,2]
  params_bm_NA[i, ] <- c(SigmaS,Sd)
  
  X_bm_NA[i, ] <- traj_sim
  
}

temps<-simRecords(dMor_0,
                         N=N,
                         gaussConv = TRUE,
                         sigmaGauss = 1,
                         sigmaSmooth=2.8,
                         sd = 78.6)[,1]
```

### e) Enregistrement des données simulées et des paramètres associés

```{r}
#save(temps,X_bm_NA,params_bm_NA, file = "train_samples_NBIVF_big_bossons_NA_nm.RData")
data<-load(file = "train_samples_NBIVF_big_bossons_NA_nm.RData")
print(data)
```

### f) traitement des données

```{r}
# Normalisation légère de X 
C <-max(abs(X_bm_NA))
X_scaled <- as.matrix(X_bm_NA /C)

# Normalisation des paramètres (Y)
params_scale <- scale(params_bm_NA)
Y_scaled <- as.matrix(params_scale)

# Split train/test
set.seed(42)
n_total <- nrow(X_scaled)
test_frac <- 0.2
test_idx <- sample(1:n_total, size = floor(test_frac * n_total))
train_idx <- setdiff(1:n_total, test_idx)

X_train <- X_scaled[train_idx, ]
Y_train <- Y_scaled[train_idx, ]

X_test  <- X_scaled[test_idx, ]
Y_test  <- Y_scaled[test_idx, ]

set.seed(123)
missing_idx <- which(runif(length(X_train)) < 0.3)  # 30% de trous
X_train[missing_idx] <- NA
mask_train <- ifelse(is.na(X_train), 0, 1)
mask_test  <- ifelse(is.na(X_test), 0, 1)

# Remplace les NAs par 0
X_train_filled <- ifelse(is.na(X_train), 0, X_train)
X_test_filled  <- ifelse(is.na(X_test), 0, X_test)

```

# III - Définition du modèle

## 1 - Architecture

### a) Définition de l'architecture

```{r}
input_x <- layer_input(shape = c(ncol(X_train_filled)), name = "traj_input")
input_m <- layer_input(shape = c(ncol(mask_train)), name = "mask_input")

concat <- layer_concatenate(list(input_x, input_m))

hidden <- concat %>%
  layer_dense(64, activation = "softplus",
              kernel_initializer = initializer_he_normal()) %>%
  layer_dense(32, activation = "softplus",
              kernel_initializer = initializer_he_normal())

y_out <- hidden %>%
  layer_dense(ncol(Y_train), activation = "linear")

model_BM <- keras_model(inputs = list(input_x, input_m), outputs = y_out)
```

### b) Possibilité d'avoir des données manquantes

```{r}
masked_mse <- function(y_true, y_pred, mask) {
  K <- backend()
  mask <- K$cast(mask, K$floatx())
  diff <- K$square(y_true - y_pred)
  masked_diff <- diff * mask
  loss <- K$sum(masked_diff) / (K$sum(mask) + 1e-8)
  return(loss)
}

masked_loss_wrapper <- function(mask_input) {
  function(y_true, y_pred) {
    K <- backend()
    diff <- K$square(y_true - y_pred)
    masked_diff <- diff * mask_input
    loss <- K$sum(masked_diff) / (K$sum(mask_input) + 1e-8)
    return(loss)
  }
}
```

## 2 - Compilation

```{r, results='hide'}
model_BM$compile(
  optimizer = optimizer_adam(learning_rate = 0.001),
  loss = "mse", 
  #loss=masked_loss_wrapper(mask_input),
  metrics = list(keras$metrics$MeanSquaredError())
)
```

## 3 - Entraînement

### a) Si on veut un early stop

```{r, results='hide'}
early_stop <- callback_early_stopping(
  monitor = "val_loss",
  patience = 10,
  restore_best_weights = TRUE
)
```

### b) Entraînement

```{r, eval=FALSE}
history <- model_BM$fit(
  x = list(X_train_filled, mask_train),
  y = Y_train,
  epochs = 100L,
  batch_size = 32L,
  verbose = 0,
  validation_split = 0.2,
  #callbacks = list(early_stop)
)
```

### c) Sauvegarde des données d'entraînement et du modèle

```{r}
#h<-history$history$loss
#h_val<-history$history$val_loss
#save(h,h_val, file = "history_NBIVF_big_bossons_NA_nm.RData")
data<-load(file = "history_NBIVF_big_bossons_NA_nm.RData")
print(data)

# Sauvegarde complète du modèle
#save_model_hdf5(model_BM, "mon_modele.h5")

# Rechargement plus tard
model_BM <- load_model_hdf5("mon_modele.h5",compile=FALSE)
```

# IV - Résultats du modèle

## 1 - Evolution de la fonction de perte

```{r}
plot(h, type="l", col="blue", lwd=2,xlab='epoch',ylab='loss')
lines(h_val, col="red", lwd=2)
legend("topright", legend=c("Train","Val"), col=c("blue","red"), lty=1, lwd=2)
```

## 2 - Prédiction sur les échantillons de test

```{r,results='hide'}
pred_scaled <- model_BM$predict(list(X_test_filled, mask_test),verbose=0)


mu <- attr(params_scale, "scaled:center")
sd <- attr(params_scale, "scaled:scale")

pred <- sweep(pred_scaled, 2, sd, `*`) # dénormalisation
pred <- sweep(pred, 2, mu, `+`)
```

## 3 - Quantification des prédictions

### a)  Definitions des métriques

```{r}
rmse <- function(y_true, y_pred) {
  sqrt(mean((y_true - y_pred)^2))
}

mae <- function(y_true, y_pred) {
  mean(abs(y_true - y_pred))
}

r2 <- function(y_true, y_pred) {
  1 - sum((y_true - y_pred)^2) / sum((y_true - mean(y_true))^2)
}
```

### b) Calcul pour chaque paramètre

```{r}

rmse_sigmaS <- rmse(params_bm_NA[test_idx,1], pred[,1])
rmse_Sd     <- rmse(params_bm_NA[test_idx,2], pred[,2])

mae_sigmaS <- mae(params_bm_NA[test_idx,1], pred[,1])
mae_Sd     <- mae(params_bm_NA[test_idx,2], pred[,2])

r2_sigmaS <- r2(params_bm_NA[test_idx,1], pred[,1])
r2_Sd     <- r2(params_bm_NA[test_idx,2], pred[,2])

# Affichage
cat("RMSE SigmaS :", rmse_sigmaS, "\n")
cat("RMSE Sd     :", rmse_Sd, "\n")
cat("MAE SigmaS  :", mae_sigmaS, "\n")
cat("MAE Sd      :", mae_Sd, "\n")
cat("R² SigmaS   :", r2_sigmaS, "\n")
cat("R² Sd       :", r2_Sd, "\n")

```

### c) Données manquantes

```{r}

mask_real <- ifelse(is.na( X_NA[mor1_0:morf_0]), 0, 1)
X_real_filled <- ifelse(is.na(X_NA[mor1_0:morf_0]), 0,X_NA[mor1_0:morf_0]/C)
X_real_filled <- matrix(X_real_filled, nrow = 1)
mask_real     <- matrix(mask_real, nrow = 1)

pred_scaled_bo <- model_BM$predict(list(X_real_filled, mask_real),verbose=0)

pred_bo <- sweep(pred_scaled_bo, 2, sd, `*`) # dénormalisation
pred_bo <- sweep(pred_bo, 2, mu, `+`)

print(pred_bo)
best_Ssm<-pred_bo[,1]
best_Sd<-pred_bo[,2]
```

## 4 - Représentation graphique

### a) Scatter plots et distribution des erreurs
```{r,results='hide'}
# Préparer les données pour ggplot
df_sigmaS <- data.frame(
  réel = params_bm_NA[test_idx,1],
  prédit = pred[,1],
  erreur = abs(params_bm_NA[test_idx,1] - pred[,1])/params_bm_NA[test_idx,1]
)

quant_cut <- quantile(df_sigmaS$erreur, 0.95, na.rm = TRUE)
df_sigmaS_filtre <- df_sigmaS[df_sigmaS$erreur < quant_cut, ]


df_Sd <- data.frame(
  réel = params_bm_NA[test_idx,2],
  prédit = pred[,2],
  erreur = abs(params_bm_NA[test_idx,2] - pred[,2])/params_bm_NA[test_idx,2]
)

# Scatter plot réel vs prédit


p1 <- ggplot(df_sigmaS, aes(x=réel, y=prédit)) +
  geom_point(color=alpha('#8F09C8',0.1), alpha=0.1) +
  geom_abline(intercept=0, slope=1, color="red", linetype="dashed", size=0.5) +
  annotate("text", x =8, 
           y = 0, 
           label = "--- : x = y", color = "red", size = 4, fontface = "italic") +
  xlab("Real s") + ylab("Prediction") +
  ggtitle("Scatter plot for s") +
  theme_minimal()

p2 <- ggplot(df_Sd, aes(x=réel, y=prédit)) +
  geom_point(color=alpha('#D1C700',0.1), alpha=0.1) +
  geom_abline(intercept=0, 
              slope=1, 
              color="red", 
              linetype="dashed", 
              size=0.5) +
  ggtitle("Scatter plot for sigma") +
  annotate("text", x =175, 
           y = 20, 
           label = "--- : x = y", color = "red", size = 4, fontface = "italic") +
  xlab("Real sigma") + ylab("Prediction") +
  theme_minimal()

# Histogrammes des erreurs
p3 <- ggplot(df_sigmaS_filtre, aes(x=erreur)) +
  geom_histogram( bins=20,fill=alpha('#8F09C8',0.5), color="white", alpha=0.7) +
  ggtitle("MAPE for s") + xlab("MAPE") + ylab("Count") +
  theme_minimal()

p4 <- ggplot(df_Sd, aes(x=erreur)) +
  geom_histogram(bins=20, fill=alpha('#D1C700',0.5), color="white", alpha=0.7) +
  ggtitle("MAPE for sigma") + xlab("MAPE") + ylab("Count") +
  theme_minimal()

# Combiner les plots dans une grille 2x2
grid.arrange(p1, p2, p3, p4, ncol=2)
```

# V - Comparaison entre paramètres prédits et paramètres aléatoires
## 1 - Création d'échantillons
```{r,eval=FALSE}
# Paramètres prédits
best_s<-2.8
best_sigma<-78.6
ss_pb<-best_s
sd_pb<-best_sigma
n_test_pb<-10000

# Matrice
X_pb_NA<-matrix(nrow=n_test_pb,ncol=N)

# Simulation
for (i in 1:n_test_pb) {
  #print(i)  

  traj_sim <- simRecords(dMor_0,
                         N=N,
                         gaussConv = TRUE,
                         sigmaGauss = SigmaG,
                         sigmaSmooth=ss_pb,
                         sd = sd_pb)[,2]
  
  X_pb_NA[i, ] <- traj_sim
  
}
```

```{r}
#save(X_pb_NA, file = "X_pb_NA_NBIVF_big_bossons_NA_nm.RData")
data<-load(file = "X_pb_NA_NBIVF_big_bossons_NA_nm.RData")
print(data)
```

## 2 - Comparaisons
### a) Moyennes et enveloppes
```{r}
moyenne_all <- apply(X_bm_NA, 2, mean)
q5_all <- apply(X_bm_NA, 2, quantile, probs = 0.05)
q95_all <- apply(X_bm_NA, 2, quantile, probs = 0.95)


moyenne_pb <- apply(X_pb_NA, 2, mean)
q5_pb <- apply(X_pb_NA, 2, quantile, probs = 0.05)
q95_pb <- apply(X_pb_NA, 2, quantile, probs = 0.95)

```

### b) Aire des enveloppes

```{r}
aire_pb<- sum((q95_pb - q5_pb)) * mean(diff(temps))
print(aire_pb)
aire_all<- sum((q95_all - q5_all)) * mean(diff(temps))
print(aire_all)
```


### c) Figure 10
Area and mean of the simulated trajectories with or without the predicted parameters. The black line is the real trajectory of the \textit{Glacier des Bossons}. Blue elements correspond to simulations made with random parameters $\sigma$ and $s$ (chosen according to their prior distribution). Orange elements correspond to simulations made with the parameters predicted by the neural network (presented in table \ref{tab:results_param_bossons}). Transparent areas are the envelopes (see equations \eqref{eq:quantile} and \eqref{eq:fct_dist})  of the simulations and the dotted lines are the mean of the simulations.

```{r}

plot(temps,moyenne_all,col='#6190E8',lwd=2,lty=2,xlab = "Time (year)", ylab = "Position (m)",ylim=c(min(q5_all),max(moyenne_pb)),type='l')
lines(temps,moyenne_pb,col='#E8BB61',lwd=2,lty=2)

polygon(c(temps, rev(temps)), c(q5_pb, rev(q95_pb)),
        col = alpha('orange',0.2), border = NA)
polygon(c(temps, rev(temps)), c(q5_all, rev(q95_all)),
        col = alpha('blue',0.1), border = NA)
lines((mor1_0:morf_0),X_NA[(mor1_0:morf_0)],type='l',lwd=3)
points(dMor_0[,1],dMor_0[,2],pch=19,cex=2,col='blue')
legend(
  "bottomright",
  legend = c(expression(bold("With predictions")), "90% interval", "Mean",expression(bold("Without predictions")), "90% interval", "Mean"),
  col = c(NA,  alpha("orange", 0.2), "#E8BB61",NA, alpha("blue", 0.1), "#6190E8"),
  lwd = c(2, 10, 2,2,  10, 2),
  lty = c(1, 2,2, 1, 2,2),
  bty = "n"
)


```

### d) Calcul des meilleures trajectoires

```{r}
mse <- function(a,b) mean((a-b)^2)
MAPE <- function(a, b) {
  mean(abs(a - b) / abs(b), na.rm = TRUE)
}
w<- which(!is.na(X_NA[mor1_0:morf_0]))

err_pred <- apply(X_pb_NA[,w], 1,function(x) mse((X_NA[mor1_0:morf_0])[w]  , x))
err_rand <-apply(X_bm_NA[,w], 1,function(x) mse((X_NA[mor1_0:morf_0])[w], x))

w1<-which(err_pred==min(err_pred))
w2<-which(err_rand==min(err_rand))
err_pred_MAPE <- apply(X_pb_NA[,w],1, function(x) MAPE((X_NA[mor1_0:morf_0])[w], x))
err_rand_MAPE <- apply(X_bm_NA[,w], 1,function(x) MAPE((X_NA[mor1_0:morf_0])[w], x))
```

### e) Figure 11
Simulated trajectories with or without the predicted parameters. The black line is the real trajectory of the \textit{Glacier des Bossons}. Blue elements correspond to simulations made with random parameters $\sigma$ and $s$ (chosen according to their distribution). Orange elements correspond to simulations made with the parameters predicted by the neural network (presented in table \ref{tab:results_param_bossons}). Ten random trajectories for each parameters area have been represented as well as the best trajectory in term of MSE (see Equation \ref{mse}).
```{r}

plot(temps,X_pb_NA[w1,],col='orange',lwd=3,type='l',
     xlab = "Time (year)", ylab = "Position (m)",ylim=c(min(q5_all),max(moyenne_pb)))
lines(temps,X_bm_NA[w2,],col='blue',lwd=3)
for(i in 1:5){
  lines(temps,X_bm_NA[i,],col='#54B8D6',lwd=1,lty=2)
  lines(temps,X_pb_NA[i,],col='#EDBC80',lwd=1,lty=2)
}
lines(temps,X_pb_NA[w1,],col='orange',lwd=2,type='l',
     xlab = "Time (year)", ylab = "Position (m)",
     main = "Brownian method with the best predicted parameters",ylim=c(min(q5_all),max(moyenne_pb)))
lines(temps,X_bm_NA[w2,],col='blue',lwd=2)
#polygon(c(temps, rev(temps)), c(q5_pb, rev(q95_pb)),
 #       col = alpha('orange',0.2), border = NA)
#polygon(c(temps, rev(temps)), c(q5_all, rev(q95_all)),
 #       col = alpha('blue',0.1), border = NA)


legend(
  "bottomright",
  legend = c(expression(bold("With predictions : ")), "Best trajectory","5 other trajectories", expression(bold("Without predictions : ")), "Best trajectory","5 other trajectories"),
  col = c(NA, "orange",'#EDBC80', NA, "blue",'#54B8D6'),
  lwd = c(3, 3,1,3, 3,1),
  lty = c(1, 1,2, 1,1,2),
  bty = "n"
)
lines((mor1_0:morf_0),X_NA[(mor1_0:morf_0)],type='l',lwd=3)
points(dMor_0[,1],dMor_0[,2],pch=19,cex=2,col='blue')
legend("bottomleft",legend=c('Moraines','Real trajectory'),lty=c(NA,1),lwd=c(NA,3),col=c('blue',1),pch=c(19,NA),bty = "n")
```

### f) Résumé des erreurs

```{r}
summary_pred <- c(mean = mean(err_pred), median = median(err_pred), sd = sd(err_pred))
summary_rand <- c(mean = mean(err_rand), median = median(err_rand), sd = sd(err_rand))
print("Résumé MSE (prédit vs aléatoire) :")
print(round(rbind(pred = summary_pred, rand = summary_rand), 5))

summary_pred <- c(mean = mean(err_pred_MAPE), median = median(err_pred_MAPE), sd = sd(err_pred_MAPE))
summary_rand <- c(mean = mean(err_rand_MAPE), median = median(err_rand_MAPE), sd = sd(err_rand_MAPE))
print("Résumé MAPE (prédit vs aléatoire) :")
print(round(rbind(pred = summary_pred, rand = summary_rand), 5))
print(mean(err_pred_MAPE)/mean(err_rand_MAPE))

```

### g) Figure 8
Violin plots of the distributions of the 95\% interval of the errors with the predicted parameters or with random ones. Subfigure a) illustrates the MSE  (see Equation \eqref{mse}) and subfigure b) the MAPE (see Equation \eqref{mape}). For both subfigures, the violin plots show the distribution of the error between the real \textit{Glacier des Bossons} trajectory and the predictions. The left violin plots are made on simulations made with random parameters $\sigma$ and $s$ (chosen according to their prior distribution) and the right violin plots are made on simulations made with the parameters predicted by the neural network (presented in Table \ref{tab:results_param_bossons}).  

```{r}

# Fusionner les deux ensembles d'erreurs dans un seul data frame
df_errors <- data.frame(Predicted=err_pred,Random=err_rand)
# Transformation au format "long"
df_long <- df_errors |>
  pivot_longer(cols = everything(),
               names_to = "type",
               values_to = "MSE")

# (Optionnel) filtrage des extrêmes si tu veux couper le 95e percentile
quant_cut <- quantile(df_long$MSE, 0.95, na.rm = TRUE)
df_long <- df_long[df_long$MSE < quant_cut, ]

ggplot(df_long, aes(x = type, y = MSE, fill = type)) +
  geom_violin(trim = FALSE, alpha = 0.4, linewidth = 0) +
  geom_boxplot(width = 0.08, fill = "white", outlier.shape = NA) +
  scale_fill_manual(values = c(
    "Predicted" = "orange",   # rouge
    "Random" = "blue"       # jaune
  )) +
  xlab("Parameters used") + 
  ylab("MSE") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
```

```{r}

# Fusionner les deux ensembles d'erreurs dans un seul data frame
df_errors <- data.frame(Predicted=err_pred_MAPE,Random=err_rand_MAPE)
# Transformation au format "long"
df_long <- df_errors |>
  pivot_longer(cols = everything(),
               names_to = "type",
               values_to = "MAPE")

quant_cut <- quantile(df_long$MAPE, 0.90, na.rm = TRUE)
df_long <- df_long[df_long$MAPE < quant_cut, ]



ggplot(df_long, aes(x = type, y = MAPE, fill = type)) +
  geom_violin(trim = FALSE, alpha = 0.4, linewidth = 0) +
  scale_fill_manual(values = c(
    "Predicted" = "orange",   # rouge
    "Random" = "blue"       # jaune
  )) +
  geom_boxplot(width = 0.08, fill ='white', outlier.shape = NA) +
  xlab("Parameters used") + ylab("MAPE") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")



```



```{r}
# stat sur les erreurs en fct des paramètres 
R2_s<-c(0.967,0.959,0.965,0.969,0.966,0.953,0.970)
R2_s_frac<-R2_s/sum(R2_s)
R2_sigma<-c(0.796,0.795,0.739,0.778,0.795,0.667,0.764)
R2_sigma_frac<-R2_sigma/sum(R2_sigma)
R2_mean<-(R2_s+R2_sigma)/2
pred_s<-c(3.132,2.014,2.734,2.883,2.816,2.826,3.478)
pred_sigma<-c(83.982,78.749,64.597,77.527,76.004,85.890,83.167)
aire<-c(61830,50149,58502,57782,64683,59509)
aire_frac<-aire/104407
err<-c(19083,16925,17996,17862,20498,18055)
err_frac<-err/45899

best_s<-0
best_sigma<-0
for(i in 1:length(R2_s)){
  best_s<-best_s+R2_s_frac[i]*pred_s[i]
  best_sigma<-best_sigma+R2_sigma_frac[i]*pred_sigma[i]
}
print(best_s)
print(best_sigma)
```

# VI - Résultats approfondis
## 1 - Métriques supplémentaires pour comparer les trajectoires
### a) Définitions
```{r}
longueur_geo <- function(x, t) {
  dt <- diff(t)
  dx <- diff(x)
  sum(sqrt(dt^2 + dx^2))
}

longueur_verticale <- function(x) {
  sum(abs(diff(x)))
}

variance_temporelle <- function(x) {
  var(x)
}

variance_increments <- function(x) {
  var(diff(x))
}

energie_quad <- function(x) {
  mean(diff(x)^2)
}

courbure_moyenne <- function(x) {
  mean(abs(diff(x, differences = 2)))
}

aire_sous_courbe <- function(x, t) {
  sum( (x[-1] + x[-length(x)]) / 2 * diff(t) )
}
```

### b) Ce qu'on compare

```{r}
traj1<-X_bm_NA
traj2<-X_pb_NA
t<-1:(length(traj1[1,]))
```

### c) Calcul des métriques

```{r}
L1 <- apply(traj1, 1, longueur_geo, t = t)
L2 <- apply(traj2, 1, longueur_geo, t = t)

LV1 <- apply(traj1, 1, longueur_verticale)
LV2 <- apply(traj2, 1, longueur_verticale)

E1 <- apply(traj1, 1, energie_quad)
E2 <- apply(traj2, 1,  energie_quad)

C1 <- apply(traj1, 1, courbure_moyenne)
C2 <- apply(traj2, 1, courbure_moyenne)

V1 <- apply(traj1, 1, variance_temporelle)
V2 <- apply(traj2, 1, variance_temporelle)

VI1 <- apply(traj1, 1, variance_increments)
VI2 <- apply(traj2, 1, variance_increments)

AUC1 <- apply(traj1, 1, aire_sous_courbe,t)
AUC2 <- apply(traj2, 1, aire_sous_courbe,t)
```

### d) Summary des métriques
```{r}
summary(V1)
summary(V2)
summary(VI1)
summary(VI2)
summary(AUC1)
summary(AUC2)
summary(C1)
summary(C2)
summary(E1)
summary(E2)
summary(LV1)
summary(LV2)

```

### e) Figure 9

Violin plots of the distributions of the 95\% interval of four different metrics (first line see equations \eqref{var},\eqref{geo_length}, second line see equations \eqref{auc}, \eqref{roughness}) computed on trajectories made with the predicted parameters (orange) or with random ones (blue). Red lines represent the values of those metrics for the interpolated real trajectory.
```{r}
# AUC
df_errors <- data.frame(Predicted=AUC2,Random=AUC1)
# Transformation au format "long"
df_long <- df_errors |>
  pivot_longer(cols = everything(),
               names_to = "type",
               values_to = "Roughness")

quant_cut <- quantile(df_long$Roughness, 0.90, na.rm = TRUE)
df_long <- df_long[df_long$Roughness < quant_cut, ]



AUC<-ggplot(df_long, aes(x = type, y = Roughness, fill = type)) +
  geom_violin(trim = FALSE, alpha = 0.4, linewidth = 0) +
  scale_fill_manual(values = c(
    "Predicted" = "orange",   # rouge
    "Random" = "blue"       # jaune
  )) +
  geom_hline(yintercept = 133742.2, color = "red", linetype = "dashed", size = 1)+
  geom_boxplot(width = 0.08, fill ='white', outlier.shape = NA) +
  xlab("Parameters used") + ylab("Area Under the Curve") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")

# VI
df_errors <- data.frame(Predicted=V2,Random=V1)
# Transformation au format "long"
df_long <- df_errors |>
  pivot_longer(cols = everything(),
               names_to = "type",
               values_to = "Roughness")

quant_cut <- quantile(df_long$Roughness, 0.90, na.rm = TRUE)
df_long <- df_long[df_long$Roughness < quant_cut, ]



VI<-ggplot(df_long, aes(x = type, y = Roughness, fill = type)) +
  geom_violin(trim = FALSE, alpha = 0.4, linewidth = 0) +
  scale_fill_manual(values = c(
    "Predicted" = "orange",   # rouge
    "Random" = "blue"       # jaune
  )) +
  geom_boxplot(width = 0.08, fill ='white', outlier.shape = NA) +
  xlab("Parameters used") + ylab("Variance of increments ") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")

# E
df_errors <- data.frame(Predicted=E2,Random=E1)
# Transformation au format "long"
df_long <- df_errors |>
  pivot_longer(cols = everything(),
               names_to = "type",
               values_to = "Roughness")

quant_cut <- quantile(df_long$Roughness, 0.90, na.rm = TRUE)
df_long <- df_long[df_long$Roughness < quant_cut, ]



E<-ggplot(df_long, aes(x = type, y = Roughness, fill = type)) +
  geom_violin(trim = FALSE, alpha = 0.4, linewidth = 0) +
  scale_fill_manual(values = c(
    "Predicted" = "orange",   # rouge
    "Random" = "blue"       # jaune
  )) +
  geom_boxplot(width = 0.08, fill ='white', outlier.shape = NA) +
  geom_hline(yintercept = 1439.473, color = "red", linetype = "dashed", size = 1)+
  xlab("Parameters used") + ylab("Roughness ") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")

# L
df_errors <- data.frame(Predicted=L2,Random=L1)
# Transformation au format "long"
df_long <- df_errors |>
  pivot_longer(cols = everything(),
               names_to = "type",
               values_to = "Roughness")

quant_cut <- quantile(df_long$Roughness, 0.90, na.rm = TRUE)
df_long <- df_long[df_long$Roughness < quant_cut, ]



L<-ggplot(df_long, aes(x = type, y = Roughness, fill = type)) +
  geom_violin(trim = FALSE, alpha = 0.4, linewidth = 0) +
  scale_fill_manual(values = c(
    "Predicted" = "orange",   # rouge
    "Random" = "blue"       # jaune
  )) +
  geom_boxplot(width = 0.08, fill ='white', outlier.shape = NA) +
  geom_hline(yintercept = 5141.675, color = "red", linetype = "dashed", size = 1)+
  xlab("Parameters used") + ylab("Geometric length") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")


#LV
df_errors <- data.frame(Predicted=L2,Random=L1)
# Transformation au format "long"
df_long <- df_errors |>
  pivot_longer(cols = everything(),
               names_to = "type",
               values_to = "Roughness")

quant_cut <- quantile(df_long$Roughness, 0.90, na.rm = TRUE)
df_long <- df_long[df_long$Roughness < quant_cut, ]



LV<-ggplot(df_long, aes(x = type, y = Roughness, fill = type)) +
  geom_violin(trim = FALSE, alpha = 0.4, linewidth = 0) +
  scale_fill_manual(values = c(
    "Predicted" = "orange",   # rouge
    "Random" = "blue"       # jaune
  )) +
  geom_boxplot(width = 0.08, fill ='white', outlier.shape = NA) +
  xlab("Parameters used") + ylab("Vertical length") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")


# V
df_errors <- data.frame(Predicted=V2,Random=V1)
# Transformation au format "long"
df_long <- df_errors |>
  pivot_longer(cols = everything(),
               names_to = "type",
               values_to = "Roughness")

quant_cut <- quantile(df_long$Roughness, 0.90, na.rm = TRUE)
df_long <- df_long[df_long$Roughness < quant_cut, ]



V<-ggplot(df_long, aes(x = type, y = Roughness, fill = type)) +
  geom_violin(trim = FALSE, alpha = 0.4, linewidth = 0) +
  scale_fill_manual(values = c(
    "Predicted" = "orange",   # rouge
    "Random" = "blue"       # jaune
  )) +
  geom_boxplot(width = 0.08, fill ='white', outlier.shape = NA) +
  geom_hline(yintercept = 105776, color = "red", linetype = "dashed", size = 1)+
  xlab("Parameters used") + ylab("Temporal variance") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")


```

```{r}
library(gridExtra)

grid.arrange(V,L,AUC,E, ncol = 2)
```

